Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2016-09-06 redaction/slides_ETSG1.tex
Last ?

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the export (fas) price and the import (cif) price \vspace{0.1cm}
\item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] The export ($\simeq$ fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
\item[-] A dataset also used by Hummels (2007) (until 2004) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] At the 4-digit level on some selected years as robustness check \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}  [label=slide_method]
\frametitle{Empirical specification: The estimation strategy }
\begin{itemize}
\item With some assumptions on the specification of transport costs \& the error term, and taking logs  \hyperlink{app_method_1}{\beamergotobutton{More}}\vspace{0.1cm}
\item[$\Rightarrow$] Estimate the following equation \vspace{0.1cm}
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_with_add}
\end{equation}
\begin{itemize}
\item[-] Estimates performed using non-linear least squares \hyperlink{app_method_2}{\beamergotobutton{More}} \vspace{0.1cm}
\end{itemize}
\normalsize
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:model_with_add}) constraining $t=0$:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\normalsize
\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \hyperlink{app_method_2}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
\item[-] With only ad-valorem costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_results_summary]
\frametitle{Result 1: Estimating transport costs over time}
For both the ad-valorem and the additive components (by year \& transport mode)\vspace{0.1cm}
\begin{itemize}
\item[-] \footnotesize{Average weighted values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
\end{itemize}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline 
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\ \hline \hline 
   % \# origin country & 188 & 191 & 188 & 189 \\
   % \# products & 230 & 211 & 666 & 567 \\  
  \end{tabular}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{center}}
  \end{table}%
\begin{itemize}
\item Transport costs are sizeable \vspace{0.1cm}
\begin{itemize}
\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
\end{itemize}
\item Opening the black box of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air \& ocean resp. \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_result2]
\frametitle{Result 2: Additive transport costs do matter}
\begin{itemize}
\item International transport costs: A sizeable additive component \vspace{0.1cm}
\begin{itemize}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.1cm}
\begin{itemize}
\item[$\star$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[$\star$] An iceberg cost equal 3.2\% and 2.5\% as mean values in vessel and air respectively in presence of additive costs \vspace{0.1cm}
\item[$\star$] vs 5.8\% and 5.1\% when estimated alone \vspace{0.1cm}
\end{itemize}
\item[-] A sizeable share of the additive component in total transport costs \vspace{0.1cm}
\begin{itemize}
\item[$\star$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
\item[$\star$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1 later}} \vspace{0.1cm}
\end{itemize}
\end{itemize}
\item A result confirmed by goodness of fit comparisons \vspace{0.1cm}
\begin{itemize}
\item[-] 4 measures of goodness of fit for Models (\ref{eq:model_with_add}) (with additive TC) and (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component \hyperlink{app_goodnessfit}{\beamergotobutton{More results}} \vspace{0.1cm}
\begin{itemize}
\item[$\star$] Taking into account the additional degrees of freedom
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_fig1]
\frametitle{Result 3: Characterizing the trends of transport costs }
\begin{itemize}
\item Study the shares of both ad-valorem and additive components in total transport costs  \hyperlink{slide_result2}{\beamergotobutton{Go back to Result 2}} \hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Result 3}}
\begin{figure}[htbp]
%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
%\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Air } & {\scriptsize  (b) Vessel}\\
\includegraphics[width=5cm, height=4cm]{Fig2a_decompTC_air_3d.pdf}
& \includegraphics[width=5cm,height=4cm]{Fig2b_decompTC_vessel_3d.pdf} \\
\end{tabular}
\end{center}
\end{figure}
\begin{itemize}
\item[-] Lower overall transport costs in Air than in Vessel \vspace{0.1cm}
\item[-] Downward trend for both modes since 1974 \hyperlink{app_fig1}{\beamergotobutton{Robustness of this result}}\vspace{0.1cm}
%\item[-] Slighlty ore pronounced in ocean shipping \vspace{0.1cm}
\begin{itemize}
\item[$\star$] A 50\% decrease in Air, a 60\% decrease in Vessel \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_comment_compositioneffects]
\frametitle{Three main findings}
\begin{itemize}
\item The importance of excluding composition effects  \vspace{0.1cm}
\begin{itemize}
\item[-] The reduction of the \textit{pure} transport costs: Starts in 1985 (not in 1974) \vspace{0.1cm}
\item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}) is attributable to change in the composition of trade patterns \vspace{0.1cm}
\end{itemize}
\item Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
\item The importance of the additive component of transport costs (again) \vspace{0.1cm}
\begin{itemize}
\item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in Air transport over the 1985-2005 period \vspace{0.1cm}
\item[-] In accordance with Hummels (2007), Behar \& Venables (2011) \vspace{0.1cm}
\item[-] But... No substantial difference when additive costs are included (Panel (b))
\end{itemize}
\end{itemize}
\end{frame}

@@

\textbf{Two main possible extensions}
\begin{itemize}
\item On the empirical side: Go deeper in the structural determinants of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.1cm}
\item[-] Compare the transports costs between air and vessel (excluding composition effects) \vspace{0.1cm}
\end{itemize}
\item On the theoretical side: Use our results to explore the role of additive costs \vspace{0.1cm}
\begin{itemize}
\item[-] In shaping international trade flows (trade theory) \vspace{0.1cm}
\item[-] In affecting the international transmission of business cycles (business cycle theory)
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=app_method_2]
\frametitle{More on the estimation strategy (2)}
\begin{itemize}
\item The non-linear least squares method \vspace{0.1cm}
\begin{itemize}
\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
\item[-] The criterion for convergence: That the sum of the squares of the residuals does not not increase? from one iteration to the next \vspace{0.1cm}
\end{itemize}
\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
\item Obtaining the final estimates: More details  \vspace{0.1cm}
\begin{itemize}
\item[-] After estimating Equations (\ref{eq:model_with_add}) and (\ref{eq:model_nlI}), re-built \vspace{0.1cm}
\begin{itemize}
\item[$\star$] With additive costs:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\item[$\star$] With only iceberg costs:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item[-] Then, take the weighted average by country-product, \vspace{0.1cm}
\item[-] Using the values of each trade flow (ik-specific) over total yearly trade as a weighting scheme
\end{itemize}
\end{itemize}
\hyperlink{slide_method}{\beamergotobutton{Back to slide}}
\end{frame}

>>>>

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the export ($\simeq$ fas) price and the import (cif) price \vspace{0.1cm}
\item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export point \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
\item[-] A dataset also used by Hummels (2007) (until 2004) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] At the 4-digit level on some selected years as robustness check \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}  [label=slide_method]
\frametitle{Empirical specification: The estimation strategy }
\begin{itemize}
\item With some assumptions on the specification of transport costs \& the error term, and taking logs  \hyperlink{app_method_1}{\beamergotobutton{More}}\vspace{0.1cm}
\item[$\Rightarrow$] Estimate the following equation \vspace{0.1cm}
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_with_add}
\end{equation}
\begin{itemize}
\item[-] Estimates performed using non-linear least squares \hyperlink{app_method_2}{\beamergotobutton{More}} \vspace{0.1cm}
\end{itemize}
\normalsize
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:model_with_add}) constraining $t=0$:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\normalsize
\item  Taking the weighted average over the product-country dimension, we finally get (by year and transport mode): \hyperlink{app_method_2}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
\item[-] With only ad-valorem costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_results_summary]
\frametitle{Result 1: Estimating transport costs over time}
For both the ad-valorem and the additive components \vspace{0.1cm}
\begin{itemize}
\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
\end{itemize}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\ \hline \hline
   % \# origin country & 188 & 191 & 188 & 189 \\
   % \# products & 230 & 211 & 666 & 567 \\
  \end{tabular}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{center}}
  \end{table}%
\begin{itemize}
\item Transport costs are sizeable \vspace{0.1cm}
\begin{itemize}
\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
\end{itemize}
\item Opening the black box of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in Air \& Vessel resp. 
\item[-] Additive cost: 1.8\% and 2.9\% of the export price 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_result2]
\frametitle{Result 2: Additive transport costs do matter}
\begin{itemize}
\item International transport costs: A sizeable additive component \vspace{0.1cm}
\begin{itemize}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.1cm}
\begin{itemize}
\item[$\star$] The ad-valorem cost is reduced by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[$\star$] From 5.8\% to 3.2\% in ocean shipping (mean value over the period) \vspace{0.1cm}
\item[$\star$] From 5.1\% to 2.5\% in Air transport \vspace{0.1cm}
\end{itemize}
\item[-] A sizeable share of the additive component in total transport costs \vspace{0.1cm}
\begin{itemize}
\item[$\star$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
\item[$\star$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1 later}} \vspace{0.1cm}
\end{itemize}
\end{itemize}
\item A result confirmed by goodness of fit comparisons \vspace{0.1cm}
\begin{itemize}
\item[-] 4 measures of goodness of fit for Models (\ref{eq:model_with_add}) (with additive TC) and (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component \hyperlink{app_goodnessfit}{\beamergotobutton{More results}} \vspace{0.1cm}
\begin{itemize}
\item[$\star$] Taking into account the additional degrees of freedom
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_fig1]
\frametitle{Result 3: Characterizing the trends of transport costs }
\begin{itemize}
\item Study the shares of both ad-valorem and additive components in total transport costs  \hyperlink{slide_result2}{\beamergotobutton{Go back to Result 2}} \hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Result 3}}
\begin{figure}[htbp]
%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
%\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Air } & {\scriptsize  (b) Vessel}\\
\includegraphics[width=5cm, height=4cm]{Fig2a_decompTC_air_3d.pdf}
& \includegraphics[width=5cm,height=4cm]{Fig2b_decompTC_vessel_3d.pdf} \\
\end{tabular}\end{center}
\end{figure}
\begin{itemize}
\item[-] Lower overall transport costs in Air than in Vessel \vspace{0.1cm}
\item[-] Downward trend for both modes since 1974 \hyperlink{app_fig1}{\beamergotobutton{Robustness of this result}}\vspace{0.1cm}
%\item[-] Slighlty ore pronounced in ocean shipping \vspace{0.1cm}
\begin{itemize}
\item[$\star$] A 50\% decrease in Air, a 60\% decrease in Vessel \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=slide_comment_compositioneffects]
\frametitle{Two main findings}
\begin{itemize}
\item The importance of excluding composition effects  \vspace{0.1cm}
\begin{itemize}
\item[-] The reduction of the \textit{pure} transport costs: Starts in 1985 (not in 1974) \vspace{0.1cm}
\item[-] The reduction between 1974 and 1984 (\hyperlink{slide_fig1}{\beamergotobutton{Figure 1}}) is attributable to change in the composition of trade patterns \vspace{0.1cm}
\item[-] Overall (pure) transport costs have declined by $\simeq$ 40\% since 1985\vspace{0.1cm}
\end{itemize}
\item The importance of the additive component of transport costs (again) \vspace{0.1cm}
\begin{itemize}
\item[-] When only iceberg costs are modeled (Panel (a)): A stronger decrease in Air transport over the 1985-2005 period \vspace{0.1cm}
\item[-] In accordance with Hummels (2007), Behar \& Venables (2011) \vspace{0.1cm}
\item[-] But... No substantial difference when additive costs are included (Panel (b))
\end{itemize}
\end{itemize}
\end{frame}

@@

\textbf{Three main possible extensions}
\begin{itemize}
\item On the empirical side: \vspace{0.1cm}
\begin{itemize}
\item[(1)] Compare the trends in transports costs \textit{between} air and vessel (excluding composition effects) \vspace{0.1cm}
\item[(2)] Go deeper in the structural determinants of transport costs \vspace{0.1cm}
\begin{itemize} 
\item[$\star$] Identify the respective roles of handling costs, insurance and freight costs at the root of the import-export prices gap \vspace{0.1cm}
\end{itemize}
\end{itemize}
\item On the theoretical side: 
\begin{itemize}
\item[(3)]Use our results to explore the role of additive costs \vspace{0.1cm}
\begin{itemize}
\item[$\star$] In shaping international trade flows (trade theory) \vspace{0.1cm}
\item[$\star$] In affecting the international transmission of business cycles (business cycle theory)
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[label=app_method_2]
\frametitle{More on the estimation strategy (2)}
\begin{itemize}
\item The non-linear least squares method \vspace{0.1cm}
\begin{itemize}
\item[-] At the basis of the method: Approximate the model by a linear one and refine the parameters by successive iterations \vspace{0.1cm}
\item[-] The criterion for convergence: That the sum of the squares of the residuals does not not increase? from one iteration to the next \vspace{0.1cm}
\end{itemize}
\item Eliminate potential influence of outliers: Exclude the 5 percent of the upper and lower tails of the distribution \vspace{0.1cm}
\item Obtaining the final estimates: More details  \vspace{0.1cm}
\begin{itemize}
\item[-] After estimating Equations (\ref{eq:model_with_add}) and (\ref{eq:model_nlI}), re-built \vspace{0.1cm}
\begin{itemize}
\item[$\star$] With additive costs:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\item[$\star$] With only iceberg costs:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item[-] Then, take the weighted average by country-product, \vspace{0.1cm}
\item[-] Using the values of each trade flow ($ik$-specific) over total yearly trade as a weighting scheme
\end{itemize}
\end{itemize}
\hyperlink{slide_method}{\beamergotobutton{Back to slide}}
\end{frame}
